---
title: "AIS 2011-2013 Density Analysis: Bering Strait Subset""
author: "Kelly Kapsar"
date: "March 6, 2018"
output: html_document
---

Import libraries. 

```{r include=FALSE}
library(knitr)
library(raster)
library(sp)
library(rgdal)
library(raster)
library(RArcInfo)
library(tmap)
library(tmaptools)
library(nPacMaps)
library(tidyverse)
library(sf)
library(kableExtra)
library(rgeos)

```

Relevant functions. 

```{r}
###################### Multiplot function ###################### 
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

Location to save plots. 

```{r}
path <- "C:/Users/Kelly Kapsar/Documents/MSU/Conferences/USIALE 2018/Figures/"
```

Upload merged ais data file with vessel attributes and sea ice concentration raster file. 

```{r}
ais <- read_sf("C:/Users/Kelly Kapsar/Documents/MSU/Data/SNAP_AIS/WorkingFiles/ais.shp")
icecon <- raster("C:/Users/Kelly Kapsar/Documents/MSU/Data/NSIDC_GO2135_20180218/N_20100101_concentration_v3.0.tif")
#ais.bssub <- read_sf("C:/Users/Kelly Kapsar/Documents/MSU/Data/SNAP_AIS/WorkingFiles/aisbssub.shp")
EPSG <- make_EPSG()
```

Ensure all data have the correct datum and projection. 

```{r}
#AIS DATA CRS == epsg:6393 NAD83(2011) / Alaska Albers  "+init=epsg:6393"
AKAlbers <- "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs"

#Assure the same crs for all data
ais <- st_transform(ais, AKAlbers)
extent <- c(-3575120, -1190480, 699280, 3775120)
icecon_crop <- crop(icecon, extent)
icecon_proj <- projectRaster(icecon_crop, crs = AKAlbers)
```

Create a subset of all the vessel track lines located within the Bering Strait region. 

```{r}
#Create the clipping polygon for the study area
#Lat/lon boundaries = 172, -157, 60, 66
coords <- list(matrix(c(172, 60, 172, 66, -157, 66, -157, 60, 172, 60), ncol=2, byrow=TRUE))
proj <- "+proj=latlong +datum=NAD83 +ellps=GRS80"
CP <- st_sfc(st_polygon(coords), crs = proj)
CP <- st_transform(CP, AKAlbers)

#Crop vessel track lines to the extent of the study area
ais.bssub <- st_intersection(ais, CP)

```

Create full and cropped basemaps for vessel traffic. 

```{r}
BBmap <- nPacMaps::bering()
BBmap <- st_transform(BBmap, AKAlbers)


crop_custom <- function(poly.sf, cropper) {
  poly.sp <- as(poly.sf, "Spatial")
  poly.sp <- gBuffer(poly.sp, byid=TRUE, width=0)
  cropper.sp <- as(cropper, "Spatial")
  poly.sp.crop <- crop(poly.sp, extent(cropper.sp))
  st_as_sf(poly.sp.crop)
}

CP.buff <- st_as_sf(gBuffer(as(CP, "Spatial"), width=1000))

BBmap.crop <- crop_custom(BBmap, CP)

```

Create a grid over marine space in the study area. 

```{r}
#Create a grid over the study area
CP_grid <- st_make_grid(CP, cellsize=25000, #I think this cell size is 25 sq skm 
                                 crs = AKAlbers, what = 'polygons') %>%
  st_sf('geometry' = ., data.frame('ID' = 1:length(.)))


#Remove land pixels from ice concentration raster 
land.mask <- mask(icecon_proj, icecon_proj < 1000, maskvalue=FALSE)

#Create a list with the average ice concentration value for all cells in the CP grid
#(Do this because one CP grid cell overlaps several different ice concentration raster pixels)
ice <- c()
for(i in 1:length(CP_grid$geometry)){
  ice_val <- raster::extract(land.mask, as_Spatial(CP_grid$geometry[i]))
  ifelse(anyNA(ice_val[[1]]), ice <- c(ice, NA), ice <- c(ice, mean(ice_val[[1]])))
} 
land.mask
#Convert to percentages (for some reason they're originally out of 1000?)
ice <- ice/10

#Add that list as an attribute of the CP grid
CP_grid$ice <- ice

#NA values indicate pixels with land in them 
land.pix <- which(is.na(CP_grid$ice))
CP_grid <- CP_grid[-land.pix,]

#Remove pixels outside of the study area 
out <- c()
for(i in 1:length(CP_grid$geometry)){
  if(st_contains(CP, CP_grid$geometry[i], sparse=F)==F){
    out <- c(out, i)
  }
}
CP_grid <- CP_grid[-out,]

```

Vessel traffic intensity calculations. 

```{r, include=FALSE}

#Pre-set intensity values 
CP_grid$intensity.all <- -999

for(i in 1:length(CP_grid$intensity.all)){
    #print(i)
    c <- st_intersection(ais.bssub, CP_grid$geometry[i])
    CP_grid$intensity.all[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

```

Categorize vessels into functional categories (sending/receiving/spillover/local). 

```{r}

#Collect starting/ending coordinates for each line
start.pts <- data.frame(lat = ais.bssub$lat_bgn, lon = ais.bssub$lon_bgn)
start.dat <- st_as_sf(start.pts, coords = c("lon", "lat"), crs = "+proj=longlat")
start.dat <- st_transform(start.dat, st_crs(ais.bssub)) 

end.pts <- data.frame(lat = ais.bssub$lat_end, lon = ais.bssub$lon_end)
end.dat <- st_as_sf(end.pts, coords = c("lon", "lat"), crs= "+proj=longlat")
end.dat <- st_transform(end.dat, st_crs(ais.bssub))

all.pts <- rbind(start.pts, end.pts)
all.dat <- st_as_sf(all.pts, coords = c("lon", "lat"), crs = "+proj=longlat")
all.dat <- st_transform(all.dat, st_crs(ais.bssub)) 


#Tells whether starting and ending points are inside or outside of the study area 
tele <- c()
for(i in 1:length(ais.bssub$geometry)){
  #print(i)
  ifelse(st_contains(CP, start.dat$geometry[i], sparse=F) & 
           st_contains(CP, end.dat$geometry[i], sparse=F), 
    tele <- c(tele,"intra"), 
    ifelse(st_contains(CP, start.dat$geometry[i], sparse=F) & 
             st_contains(CP, end.dat$geometry[i], sparse=F)==F, 
      tele <- c(tele,"send"), 
      ifelse(st_contains(CP, start.dat$geometry[i], sparse=F)==F & 
               st_contains(CP, end.dat$geometry[i], sparse=F),
        tele <- c(tele,"receive"), 
        ifelse(st_contains(CP, start.dat$geometry[i], sparse=F)==F & 
                 st_contains(CP, end.dat$geometry[i], sparse=F)==F,
          tele <- c(tele, "trans"), tele <- c(tele, "WTF")))))
}
ais.bssub$tele <- tele

```

Calculate the proportion of vessels in 2011 and 2012 from each vessel type and functional category. 

```{r}

#Remove years with incomplete data (.mid indicates only 2011 and 2012 track lines)
ais.bssub.mid <- ais.bssub[which(ais.bssub$year != 2010 & ais.bssub$year != 2013),]


#Calculate intensity for just 2011 and 2012 (removing incomplete years)
CP_grid$intensity.mid <- -999 #.mid = for only 2011 and 2012 years 

for(i in 1:length(CP_grid$intensity.mid)){
    #print(i)
    c <- st_intersection(ais.bssub.mid, CP_grid$geometry[i])
    CP_grid$intensity[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

#Subset data by functional type 
intra <- ais.bssub[which(ais.bssub$tele == "intra"),]
send <- ais.bssub[which(ais.bssub$tele == "send"),]
receive <- ais.bssub[which(ais.bssub$tele == "receive"),]
trans <- ais.bssub[which(ais.bssub$tele == "trans"),]

intra.mid <- ais.bssub.mid[which(ais.bssub.mid$tele == "intra"),]
send.mid <- ais.bssub.mid[which(ais.bssub.mid$tele == "send"),]
receive.mid <- ais.bssub.mid[which(ais.bssub.mid$tele == "receive"),]
trans.mid <- ais.bssub.mid[which(ais.bssub.mid$tele == "trans"),]

#Count frequency of vessels for each functional type 
types <- c(unique(ais.bssub.mid$ShptyL2))
func <- c("intra", "send", "receive", "trans")

#Need to figure out how to pre-specify column names in the dataframe 

df <- setNames(data.frame(matrix(ncol = length(unique(send$ShptyL2)), nrow = length(func))), c(unique(send$ShptyL2)))
row.names(df) <- c(func)

#Couldn't figure out the code-efficient way to populate the dataframe, so I'm just going to spell it out

#One way I tried to loop through and generate the values
#for(f in func){ 
#  for(i in types){
#    l <- length(f$ShptyL2[which(f$ShptyL2 == i)])
#    df[f, i] <- l
#  }
#}

#a second way was to aggregate them and then append
#df.ag <- aggregate(x = send, by = list(unique.values = send$ShptyL2), FUN = length)

#Manual, brute force way
df["send","Fishing"] <- length(send.mid$ShptyL2[which(send.mid$ShptyL2 == "Fishing")])
df["send","Offshore"] <- length(send.mid$ShptyL2[which(send.mid$ShptyL2 == "Offshore")])
df["send","Tankers"] <- length(send.mid$ShptyL2[which(send.mid$ShptyL2 == "Tankers")])
df["send","Miscellaneous"] <- length(send.mid$ShptyL2[which(send.mid$ShptyL2 == "Miscellaneous")])
df["send","Bulk Carriers"] <- length(send.mid$ShptyL2[which(send.mid$ShptyL2 == "Bulk Carriers")])
df["send","Dry Cargo/Passenger"] <- length(send.mid$ShptyL2[which(send.mid$ShptyL2 == "Dry Cargo/Passenger")])
df["send","Non-Merchant Ships"] <- length(send.mid$ShptyL2[which(send.mid$ShptyL2 == "Non-Merchant Ships")])

df["receive","Fishing"] <- length(receive.mid$ShptyL2[which(receive.mid$ShptyL2 == "Fishing")])
df["receive","Offshore"] <- length(receive.mid$ShptyL2[which(receive.mid$ShptyL2 == "Offshore")])
df["receive","Tankers"] <- length(receive.mid$ShptyL2[which(receive.mid$ShptyL2 == "Tankers")])
df["receive","Miscellaneous"] <- length(receive.mid$ShptyL2[which(receive.mid$ShptyL2 == "Miscellaneous")])
df["receive","Bulk Carriers"] <- length(receive.mid$ShptyL2[which(receive.mid$ShptyL2 == "Bulk Carriers")])
df["receive","Dry Cargo/Passenger"] <- length(receive.mid$ShptyL2[which(receive.mid$ShptyL2 == "Dry Cargo/Passenger")])
df["receive","Non-Merchant Ships"] <- length(receive.mid$ShptyL2[which(receive.mid$ShptyL2 == "Non-Merchant Ships")])

df["trans","Fishing"] <- length(trans.mid$ShptyL2[which(trans.mid$ShptyL2 == "Fishing")])
df["trans","Offshore"] <- length(trans.mid$ShptyL2[which(trans.mid$ShptyL2 == "Offshore")])
df["trans","Tankers"] <- length(trans.mid$ShptyL2[which(trans.mid$ShptyL2 == "Tankers")])
df["trans","Miscellaneous"] <- length(trans.mid$ShptyL2[which(trans.mid$ShptyL2 == "Miscellaneous")])
df["trans","Bulk Carriers"] <- length(trans.mid$ShptyL2[which(trans.mid$ShptyL2 == "Bulk Carriers")])
df["trans","Dry Cargo/Passenger"] <- length(trans.mid$ShptyL2[which(trans.mid$ShptyL2 == "Dry Cargo/Passenger")])
df["trans","Non-Merchant Ships"] <- length(trans.mid$ShptyL2[which(trans.mid$ShptyL2 == "Non-Merchant Ships")])

df["intra","Fishing"] <- length(intra.mid$ShptyL2[which(intra.mid$ShptyL2 == "Fishing")])
df["intra","Offshore"] <- length(intra.mid$ShptyL2[which(intra.mid$ShptyL2 == "Offshore")])
df["intra","Tankers"] <- length(intra.mid$ShptyL2[which(intra.mid$ShptyL2 == "Tankers")])
df["intra","Miscellaneous"] <- length(intra.mid$ShptyL2[which(intra.mid$ShptyL2 == "Miscellaneous")])
df["intra","Bulk Carriers"] <- length(intra.mid$ShptyL2[which(intra.mid$ShptyL2 == "Bulk Carriers")])
df["intra","Dry Cargo/Passenger"] <- length(intra.mid$ShptyL2[which(intra.mid$ShptyL2 == "Dry Cargo/Passenger")])
df["intra","Non-Merchant Ships"] <- length(intra.mid$ShptyL2[which(intra.mid$ShptyL2 == "Non-Merchant Ships")])

table <- kable(df, "html") %>% kable_styling()
table



df2 <- setNames(data.frame(matrix(ncol = 2, nrow = 2)), c("2011", "2012"))
row.names(df2) <- c("Total Distance Travelled (km)", "Total Number of ships")


df2["Total Distance Travelled (km)","2011"] <- sum(st_length(ais.bssub.mid[which(ais.bssub.mid$year == 2011),]))/1000
df2["Total Distance Travelled (km)","2012"] <- sum(st_length(ais.bssub.mid[which(ais.bssub.mid$year == 2012),]))/1000
df2["Total Number of ships","2011"] <- length(ais.bssub.mid$year[which(ais.bssub.mid$year == 2011)])
df2["Total Number of ships","2012"] <- length(ais.bssub.mid$year[which(ais.bssub.mid$year == 2012)])

table2 <- kable(df2, "html") %>% kable_styling()
table2
```

Calculate intensity for functional and categorical typologies to the CP Grid. 

```{r}

############################## Functional Typology ############################## 
for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(send.mid, CP_grid$geometry[i])
    CP_grid$intensity_send.mid[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(receive.mid, CP_grid$geometry[i])
    CP_grid$intensity_receive.mid[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(intra.mid, CP_grid$geometry[i])
    CP_grid$intensity_intra.mid[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(trans.mid, CP_grid$geometry[i])
    CP_grid$intensity_trans.mid[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

############################## Categorical Typology ############################## 

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(ais.bssub.mid[which(ais.bssub.mid$ShptyL2 == "Bulk Carriers"),], CP_grid$geometry[i])
    CP_grid$intensity_bulkc.mid[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(ais.bssub.mid[which(ais.bssub.mid$ShptyL2 == "Offshore"),], CP_grid$geometry[i])
    CP_grid$intensity_offshore.mid[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(ais.bssub.mid[which(ais.bssub.mid$ShptyL2 == "Tankers"),], CP_grid$geometry[i])
    CP_grid$intensity_tankers.mid[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(ais.bssub.mid[which(ais.bssub.mid$ShptyL2 == "Miscellaneous"),], CP_grid$geometry[i])
    CP_grid$intensity_miscellaneous.mid[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(ais.bssub.mid[which(ais.bssub.mid$ShptyL2 == "Fishing"),], CP_grid$geometry[i])
    CP_grid$intensity_fishing.mid[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(ais.bssub.mid[which(ais.bssub.mid$ShptyL2 == "Dry Cargo/Passenger"),], CP_grid$geometry[i])
    CP_grid$intensity_drycpass.mid[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(ais.bssub.mid[which(ais.bssub.mid$ShptyL2 == "Non-Merchant Ships"),], CP_grid$geometry[i])
    CP_grid$intensity_nonmerchant.mid[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

############################## Separated by Year ############################## 
for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(ais.bssub[which(ais.bssub$year == 2011),], CP_grid$geometry[i])
    CP_grid$intensity_2011[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(ais.bssub[which(ais.bssub$year == 2012),], CP_grid$geometry[i])
    CP_grid$intensity_2012[i] <-  sum(st_length(c$geometry)) #Total m travelled within that cell during the study period
}

############################## Proportion of Total Traffic ############################## 
#2011 proportional intensity
total2011 <- sum(st_length(ais.bssub.mid$geometry[which(ais.bssub.mid$year == 2011)]))

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(ais.bssub[which(ais.bssub$year == 2011),], CP_grid$geometry[i])
    CP_grid$prop_intensity_2011[i] <-  sum(st_length(c$geometry))/total2011 
}


sum(CP_grid$prop_intensity_2011) #Not 1 bc of land excluded pixels

#2012 proportional intensity 
total2012 <- sum(st_length(ais.bssub.mid$geometry[which(ais.bssub.mid$year == 2012)]))

for(i in 1:length(CP_grid$intensity)){
    print(i)
    c <- st_intersection(ais.bssub[which(ais.bssub$year == 2012),], CP_grid$geometry[i])
    CP_grid$prop_intensity_2012[i] <-  sum(st_length(c$geometry))/total2012 
}


sum(CP_grid$prop_intensity_2012) #Not 1 bc of land excluded pixels


#Change in proportional intensity 2011-2012

CP_grid$prop_intensity_change <- CP_grid$prop_intensity_2012 - CP_grid$prop_intensity_2011

```

Standardize intensity values to km (from m).

```{r}

CP_grid$intensity.mid.km <- CP_grid$intensity.mid/1000
CP_grid$intensity_2011.km <- CP_grid$intensity_2011/1000
CP_grid$intensity_2012.km <- CP_grid$intensity_2012/1000
CP_grid$intensity_bulkc.mid.km <- CP_grid$intensity_bulkc.mid/1000
CP_grid$intensity_drycpass.mid.km <- CP_grid$intensity_drycpass.mid/1000
CP_grid$intensity_fishing.mid.km <- CP_grid$intensity_fishing.mid/1000
CP_grid$intensity_intra.mid.km <- CP_grid$intensity_intra.mid/1000
CP_grid$intensity_miscellaneous.mid.km <- CP_grid$intensity_miscellaneous.mid/1000
CP_grid$intensity_nonmerchant.mid.km <- CP_grid$intensity_nonmerchant.mid/1000
CP_grid$intensity_offshore.mid.km <- CP_grid$intensity_receive.mid/1000
CP_grid$intensity_receive.mid.km <- CP_grid$intensity_receive.mid/1000
CP_grid$intensity_send.mid.km <- CP_grid$intensity_send.mid/1000
CP_grid$intensity_tankers.mid.km <- CP_grid$intensity_tankers.mid/1000
CP_grid$intensity_trans.mid.km <- CP_grid$intensity_trans.mid/1000

```

Save data! 

```{r}
write_sf(ais.bssub, "aisbssub.shp")
write_sf(CP_grid, "CP_grid.shp")
```

Number of ships by week of year. 

```{r}

ggplot(ais.bssub, aes(week)) + 
  geom_bar(aes(fill = year), position = "dodge") +
  labs(x="Week of year", y="Number of ships")
  
ggsave(paste(path,"WeekHistByYearAll",plt,".png", sep=""), plot=last_plot(), width=8, height=4)


ggplot(ais.bssub.mid, aes(week)) + 
  geom_bar(aes(fill = year), position = "dodge") +
  labs(x="Week of year", y="Number of ships", fill="Year")
  
ggsave(paste(path,"WeekHistByYear1112",plt,".png", sep=""), plot=last_plot(), width=8, height=4)

ggplot(ais.bssub.mid[which(ais.bssub.mid$year == 2011),], aes(week)) + 
  geom_bar(aes(fill = ShptyL2)) +
  labs(x="Week of year", y="Number of ships", fill="Ship Type")
  
ggsave(paste(path,"WeekHistByCat11",plt,".png", sep=""), plot=last_plot(), width=8, height=4)

```

Histograms for vessel intensity by functional and categorical typologies. 

```{r}

nm.func <- c("CP_grid$intensity_send.mid.km", "CP_grid$intensity_intra.mid.km", "CP_grid$intensity_receive.mid.km", "CP_grid$intensity_trans.mid.km")
xlabs.func <- c("Outbound Vessel Intensity", "Inbound Vessel Intensity","Local Vessel Intensity","Transient Vessel Intensity")

nm.cat <- c("CP_grid$intensity_bulkc.mid.km","CP_grid$intensity_drycpass.mid.km", "CP_grid$intensity_fishing.mid.km", "CP_grid$intensity_offshore.mid.km", "CP_grid$intensity_nonmerchant.mid.km", "CP_grid$intensity_tankers.mid.km", "CP_grid$intensity_miscellaneous.mid.km")
xlabs.cat <- c("Bulk Carrier Intensity", "Dry Cargo/Passenger Intensity","Fishing Vessel Intensity","Offshore Intensity", "Non-merchant Vessel Intensity", "Tanker Vessel Intensity","Miscellaneous Vessel Intensity")

#Create histograms for functional vessel intensities 
for(i in 1:length(nm.func)){
  #Convert variable for dataframe to make ggplot happy 
  dt <- data.frame(d = eval(parse(text=nm.func[i])))
  #Specify file path for graphical output
  #mypath <- file.path("C:/Users/Kelly Kapsar/Documents/MSU/Classes/Adv. Quantitative Methods in Geography/Exercise2",paste(nm[i],".png",sep=""))
  
  p <- ggplot(dt, aes(d)) +
        geom_histogram(bins = 15) +
        #ggtitle(titles[i]) +
        labs(x=xlabs.func[i], y="Number of cells")
        #theme(plot.title = element_text(hjust = 0.5))
  #ggsave(mypath, plot = last_plot())
  plt <- paste("p",i,sep="")
  assign(plt, p)
  plt
  ggsave(paste(path,"FuncHists",plt,".png", sep=""), plot=last_plot())
}


#Create histograms for categorical vessel intensities
for(i in 1:length(nm.cat)){
  #Convert variable for dataframe to make ggplot happy 
  dt <- data.frame(d = eval(parse(text=nm.cat[i])))
  #Specify file path for graphical output
  #mypath <- file.path("C:/Users/Kelly Kapsar/Documents/MSU/Classes/Adv. Quantitative Methods in Geography/Exercise2",paste(nm[i],".png",sep=""))
  
  p <- ggplot(dt, aes(d)) +
        geom_histogram(bins = 15) +
        #ggtitle(titles[i]) +
        labs(x=xlabs.cat[i], y="Number of cells")
        #theme(plot.title = element_text(hjust = 0.5))
  #ggsave(mypath, plot = last_plot())
  plt <- paste("p",i,sep="")
  assign(plt, p)
  plt
  ggsave(paste(path,"CatHists",plt,".png", sep=""), plot=last_plot())
}

```

Intensity Maps

```{r}

pos = c(0.1,0.68)

#Create intensity map for all 2011 vessels 
inten2011 <- tm_shape(BBmap.crop) + 
  tm_polygons(palette="Greys") + 
  tm_shape(CP_grid) + 
  tm_polygons(c("intensity_2011.km"), title="Vessel Traffic Intensity (km) - 2011") + 
  tm_legend(position=pos) +
  tm_scale_bar(position=c("left","bottom"))

save_tmap(inten2011, paste(path,"Inten2011.png", sep=""))


#Create intensity map for all 2012 vessels 
inten2012 <- tm_shape(BBmap.crop) + 
  tm_polygons(palette="Greys") + 
  tm_shape(CP_grid) + 
  tm_polygons(c("intensity_2012.km"), title="Vessel Traffic Intensity (km) - 2012") + 
  tm_legend(position=pos) +
  tm_scale_bar(position=c("left","bottom"))

save_tmap(inten2012, paste(path,"Inten2012.png", sep=""))


#Create intensity map for proportion of vessel traffic in each cell in 2011 
propinten2011 <- tm_shape(BBmap.crop) + 
  tm_polygons(palette="Greys") + 
  tm_shape(CP_grid) + 
  tm_polygons(c("prop_intensity_2011"), title="Vessel Traffic Intensity \n(proportional) - 2011") + 
  tm_legend(position=pos) +
  tm_scale_bar(position=c("left","bottom"))

save_tmap(propinten2011, paste(path,"propinten2011.png", sep=""))


#Create intensity map for proportion of vessel traffic in each cell in 2012 
propinten2012 <- tm_shape(BBmap.crop) + 
  tm_polygons(palette="Greys") + 
  tm_shape(CP_grid) + 
  tm_polygons(c("prop_intensity_2012"), title="Vessel Traffic Intensity \n(proportional) - 2012") + 
  tm_legend(position=pos) +
  tm_scale_bar(position=c("left","bottom"))

save_tmap(propinten2012, paste(path,"propinten2012.png", sep=""))


#Create intensity map for change in proportion of vessel traffic in each cell from 2011 to 2012 
propintenchange <- tm_shape(BBmap.crop) + 
  tm_polygons(palette="Greys") + 
  tm_shape(CP_grid) + 
  tm_polygons(c("prop_intensity_change"), title="Change in Vessel Traffic Intensity \n(proportional) \n2011- 2012") + 
  tm_legend(position=c(pos)) +
  tm_scale_bar(position=c("left","bottom"))

save_tmap(propintenchange, paste(path,"propintenchange.png", sep=""))


#Create intensity maps for all vessel categories 
for(i in 1:length(nm.cat)){
  p <- tm_shape(BBmap.crop) + 
    tm_polygons(palette="Greys") + 
    tm_shape(CP_grid) + 
    tm_polygons(substring(nm.cat[i],9), title=xlabs.cat[i]) +
    tm_legend(position=pos) +
    tm_scale_bar(position=c("left","bottom"))
  
  plt <- paste("p",i,sep="")
  assign(plt, p)
  
  save_tmap(p, paste(path,"CatIntenMaps",plt,".png", sep=""))
}


#Create intensity maps for all vessel functional categories 
for(i in 1:length(nm.func)){
  p <- tm_shape(BBmap.crop) + 
    tm_polygons(palette="Greys") + 
    tm_shape(CP_grid) + 
    tm_polygons(substring(nm.func[i],9), title=xlabs.func[i]) +
    tm_legend(position=pos) +
    tm_scale_bar(position=c("left","bottom"))
  
  plt <- paste("p",i,sep="")
  assign(plt, p)
  
  save_tmap(p, paste(path,"FuncIntenMaps",plt,".png", sep=""))
}

```


Base Maps of Study Area 

```{r}

#Base map showing study area 
CP_grid$zeros <- 0

study.area <- tm_shape(BBmap) + 
  tm_polygons(palette="Greys") + 
  tm_shape(CP_grid) +
  tm_polygons("zeros") +
  tm_legend(show=FALSE) +
  tm_scale_bar(position=c("left","bottom"))

save_tmap(study.area, paste(path,"StudyArea.png", sep=""))

#How to crop the map? 
study.area.zoom <- tm_shape(BBmap.crop) + 
  tm_polygons(palette="Greys") + 
  tm_shape(CP_grid) +
  tm_polygons("zeros") +
  tm_legend(show=FALSE) +
  tm_scale_bar(position=c("left","bottom"))

save_tmap(study.area, paste(path,"StudyAreaZoom.png", sep=""))

```

Random extra code. 

```{r}

#More plotting

plot(BeringBase)
plot(start.dat$geometry[400], col="green", pch=19, cex=3, add=T)
plot(local$geometry[400], col="purple", lwd= 3, add=T)
plot(end.dat$geometry[400], col="red", pch=19, cex=3, add=T)
plot(CP, add=T)

plot(BeringBase)
plot(ais.bssub$geometry[400], col="red", lwd= 3, add=T)
plot(ais.bssub$geometry[which(ais.bssub$MMSI_x == local$MMSI_x[400])], col="green", lwd= 3, add=T)
plot(CP, add=T)

library(rgeos)
library(ggplot2)


plot(BBmap.crop, col="darkseagreen")
plot(local$geometry[which(local$ShptyL2 == "Fishing")], add=T)

unique(local$ShptyL2)


CP_grid$intensity_2011.km

##########################################################################

#For some reason r has an issue with plotting the sf grid over a raster
#It plots the sf lines just fine over the raster, but not the grid
#It will also plot the geometry of the grid over the top, just not the 
#attributes of the grid 
plot(icecon_proj)
plot(CP_grid$geometry, add=T)

plot(CP_grid["intensity"], add=T)
plot(ais.bssub$geometry, add=T)



xyz <- c(0,1,2,NA)
anyNA(xyz)

ice_val[1]

plot(icecon_proj[is.na(values(icecon_proj))])

plot(icecon_proj, colNA="black")

plot(land.mask)
#plot(ais.bssub$geometry, add=T)
plot(CP_grid$geometry[1000], add=T)
```



